Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increases Repair Time?

### 2015-03-11 phaseDelay.tex
Outline added for Forrest

\documentclass{sig-alternate}
\usepackage{multirow}
\usepackage{color}
\usepackage{colortbl}

\usepackage{picture}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}} 
\newcommand{\lucas}[1]{\textcolor{red}{LUCAS: #1}} 
\newcommand{\bill}[1]{\textcolor{blue}{BILL: #1}} 
\newcommand{\carter}[1]{\textcolor{cyan}{CARTER: #1}} 
%\newenvironment{changed}{\par}{\par}

@@

This result raises the question:
how did myths like phase delay survive so long  without  critical reassessment.  
As a community, we need to reflect more on 
research practices and on how (and when) we reflect on the prominent hypotheses of our field.
\end{abstract}

@@

Accordingly, this paper revisits
the {\em phase delay} effect
i.e.  the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect.
Previously~\cite{me08a}, we have constructed return-on-investement models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, who many  unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers,
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier that the developers then, due to the phase delay effect, this saves the overall cost of the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

Our surveys of  industrial practitioners (shown below) reveal it to be
the single most strongly held belief amongst commercial SE engineers;
Also,
as discussed later
in this paper, concerns over phase delay were a major driver in the development of agile software
processes.

The conclusions of this paper is that it is erroneous to believe in phase delay.
When we looked for supporting evidence for that effect, we found that:
\be 
\item The evidence for that effect in the historical literature is very weak (to say the least);
\item  In  a large sample of   contemporary software projects, there is no evidence for phase delay.
 \ee
The rest of this paper present evidence for the above two points. 
The first point is  documented via  a literature survey and the second point is demonstrated
via an analysis of  
 171 software  projects developed in the period \bill{2004 to 2014}.  To the best of our knowledge,
 this is the largest study on phase delay yet seen in the literature.

The conclusions of this paper are two fold. Firstly, the SE community needs to stop
justifying changes to SE
practices via the phase delay effect. Secondly,   myths like phase delay should not have lasted decades
in the literature without  critical reassessment.  As a community, we need to reflect more on our
research practices and on how (and when) we reflect and revisit the prominent hypotheses of our field.

@@

Before exploring one particular truism, we digress to discuss the merits of revisiting
old conclusions in software engineering.

Some researchers   adamantly believe in general
principles of software
engineering. Processional societies assume such generalities exist when the offer
 lists of  supposedly general ``best practices'' such as
the IEEE 1012 standard for software verification~\cite{1012}.  Endres \& Rombach's comment that 
\begin{quote}
Based on repeated and consistent observations, key lessons of these fields can now be formulated into rules or even laws, providing initial building blocks towards a theoretical foundation that is essential for further research, for teaching and for the practice of software development~\cite{endres03,rombach11}
\end{quote}
 Endres \& Rombach offer dozens of supposedly general lessons of software engineering~\cite{rombach11}.
 Many other 
widely-cited researchers  do the same; e.g.
Glass~\cite{glass02}; Jones~\cite{jones07}; Boehm~\cite{boehm00b}.
Budgen \& Kitchenham seek to reorganize SE research using
general
conclusions drawn from a larger number of studies~\cite{kitch04,budgen09}.

@@

\item
Ray et al.~\cite{ray2014lang} tested if   strongly typed languages
predict for better code quality. In  728 projects,
they found  only a modest benefit in strong typing
(and caution that that effect may actually be due to other  conflating factors).
\item
Fenton \& Neil~\cite{fenton00,fenton00b}   critique the truism that
``pre-release fault rates for software
are a predictor for post-release failures'' (as claimed by~\cite{dunsmore88},
amongst others). For the systems described in~\cite{fenton97}, they
show that software modules that were highly fault-prone
prior to release revealed very few faults after release.
\item
Meyer claims that   object-oriented (OO) encapsulation will
reduce error rates in software~\cite{Meyer1988}.  Yet empirical results suggest
that debugging an OO program is many times harder and
longer than debugging a standard procedural program~\cite{hatton98}.
\item
A truism of visual programming is that ``visual
representations are inherently superior to mere textual representations''.
A review by Menzies suggests that the available
evidence for this claim is hardly conclusive~\cite{me00v}. 
\item Numerous recent {\em local learning} results compare single models
learned from all available data to multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,betta12,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions).
 \item 
Other work studies  found dozens of single
models that performed worse that {\em committees of models}, all with different opinions,
that contribute to an {\em ensemble} prediction of software development effort~\cite{kocaguneli2012value,azhar13}.
\ei
To be fair to the software engineering community, SE is not the only
field where practitioners hang on to persistent beliefs, even if the evidence
for those beliefs is not strong.
The medical profession applies  many practices based on studies
that have been disproved. For example,
a  recent article
in the Mayo Clinic Proceedings~\cite{prasad13} found  
146 medical practices based on studies 
in year $i$, but which were  reversed by subsequent trials within years $i+10$.
Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
Aschwanden warns:
\begin{quote}
Cognitive biases - such as motivated reasoning (all of us want to believe that the things we do make a difference), base rate neglect (failing to pay attention to what happens in the absence of the intervention), and confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) - also influence how we process information~\cite{aschwanden15}.
\end{quote}

The cognitive issues that complicate medicine are also found in software engineering.
According to Passos et al.~\cite{passos11}, commercial developers
are all too willing to believe in general
truisms.
They  warn that developers
usually assume that the truisms they learn from a few past
projects are general to 
all their future projects:
\begin{quote}\label{q:pass}
...the past experiences were taken into account without 
much consideration for their context~\cite{passos11}.  
\end{quote}
The results of Jorgensen \& Gruschke support the findings if Passos et al. They report that 
  supposed software engineering    ``experts'' rarely use lessons
  from past projects to improve their future reasoning~\cite{jorgensen09}. 
 They note that
when the experts
  fail to revise their beliefs, this leads to poor
 conclusions and software projects  (see examples in~\cite{jorgensen09}).
%% This leads to  detrimental
%% situations where groups or 
%% individuals working on the same project have conflicting beliefs, which they never debate
%or refine or unify.

In summary, just like medicine, our field suffers when
 software engineers do  revise old beliefs.  Therefore, it is important
 to regularly  reexamine    old beliefs.
\subsection{Why Study Phase Delay?}
Due to its prominence in the field of software engineering,
this paper studies phase delay. As mentioned in the introduction, concerns about phase delay have
been used for decades to motivate major changes to software engineering practices. For example,
in 2000, Kent Beck used phase delay to motivate his call for ``extreme programming'' (a set of techniques
that evolved into contemporary agile and DevOps practices). Beck writes:
\begin{quote}
One of the universal assumptions of software engineering is that the cost of changing a program rises exponentially over time. I can remember sitting in a big linoleum-floored classroom as a college junior and seeing the professor draw on the board the curve found in \fig{curve1}.

@@

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 &\rule{2mm}{2mm} \\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 &\rule{4mm}{2mm} \\
Phase $i$ &Phase $i+2$ & 1 & 4& 4 &\rule{8mm}{2mm} \\
Phase $i$ &Phase $i+3$ & 1 & 8 & 8 & \rule{16mm}{2mm} \\
 Phase $i$  &Phase $i+4$ &1 & 16 & 16 &\rule{32mm}{2mm} 
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve1}.}\label{fig:curve1a}
\end{figure}

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 &\rule{2mm}{2mm} \\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 &\rule{4mm}{2mm} \\
Phase $i$ &Phase $i+2$ & 1 & 2& 2 &\rule{4mm}{2mm} \\
Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & \rule{4mm}{2mm} \\
 Phase $i$  &Phase $i+4$ &1 & 2& 2 &\rule{4mm}{2mm} 
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
\end{figure}

@@

\section{Phase Delay and Cost-to-fix Results from Software Engineering Law Survey}

@@

\section{billTSP}

\section{carterCharts}

\input{carter}

The distribution of defects found and fixed per phase is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections.

\begin{figure*}[!t]
\begin{center}
\includegraphics{fix-phase-dist.png}
\end{center}
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure*}

@@

@@

\begin{figure*}[!t]
\begin{center}
\includegraphics[width=6in]{waterfall3.png}
\end{center}
\caption{Different approaches to software development:  waterfall and agile (bottom left).}
\end{figure*}

\section{Data}

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed\\\hline
57& 6& 16& 47&  BeforeDevelopment&Code\\
72& 2& 4& 11&  BeforeDevelopment&CodeInspect\\
165& 6& 18& 56&  BeforeDevelopment&Test\\
50& 6& 20& 37&  BeforeDevelopment&IntTest\\
42& 8& 28& 65&  BeforeDevelopment&SysTest\\\hline

66& 2& 6& 15&  Planning&Planning\\
41& 1& 5& 16&  Planning&ReqtsReview\\\hline

23& 0& 1& 3&  Reqts&Reqts\\
245& 6& 13& 23&  Reqts&ReqtsReview\\
289& 7& 16& 30&  Reqts&ReqtsInspect\\
32& 6& 21& 40&  Reqts&Design\\
49& 2& 7& 24&  Reqts&DesignInspect\\\hline

94& 2& 5& 10&  HLD&HLDReview\\
133& 3& 8& 19&  HLD&HLDInspect\\
33& 2& 5& 17&  HLD&Design\\\hline

37& 2& 4& 6&  Design&Design\\
455& 11& 24& 53&  Design&DesignInspect\\
218& 5& 14& 30&  Design&Code\\
166& 4& 11& 22&  Design&CodeInspect\\
542& 14& 31& 60&  Design&Test\\
67& 6& 15& 40&  Design&IntTest\\
92& 10& 25& 53&  Design&SysTest\\\hline

53& 2& 4& 9&  DesignInspect&DesignInspect\\
38& 1& 3& 9&  DesignInspect&Code\\\hline

126& 4& 12& 31&  Code&Code\\
459& 11& 24& 45&  Code&CodeInspect\\
461& 11& 26& 55&  Code&Test\\
348& 10& 26& 62&  Code&IntTest\\
230& 8& 18& 39&  Code&SysTest\\
71& 2& 11& 28&  Code&AcceptTest\\\hline

64& 2& 4& 10&  CodeInspect&CodeInspect\\
57& 2& 5& 13&  CodeInspect&Test\\\hline

110& 3& 10& 23&  Test&Test\\
66& 2& 4& 12&  Test&QualTest\\\hline

32& 1& 2& 4&  IntTest&QualTest\\
36& 1& 11& 30&  IntTest&IntTest\\
21& 1& 2& 5&  IntTest&SysTest\\
 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data.}
\label{fig:faw}
\end{figure}

@@

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{90th }\\
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltaBig}
\end{tabular}
\end{center}
\caption{90th percentile scale ups.}
\label{fig:scale90}
\end{figure}

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{99th }\\
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltasBigBig}
\end{tabular}
\end{center}
\caption{99th percentile scale ups.}
\label{fig:scale90}
\end{figure}

>>>>

\documentclass{sig-alternate}
\usepackage{multirow}
\usepackage[usenames, dvipsnames]{color}
\usepackage{colortbl}

\usepackage{picture}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}} 
\newcommand{\lucas}[1]{\textcolor{red}{LUCAS: #1}} \newcommand{\tim}[1]{\textcolor{orange}{TIM: #1}}
\newcommand{\bill}[1]{\textcolor{blue}{BILL: #1}} 
\newcommand{\carter}[1]{\textcolor{cyan}{CARTER: #1}} 
\newcommand{\sei}[1]{\textcolor{RedViolet}{BILL/FORREST: #1}} 
\newcommand{\todo}[1]{\textcolor{Maroon}{TODO: #1}} 
%\newenvironment{changed}{\par}{\par}

@@

%This result raises the question: how did myths like phase delay survive so long  without  critical reassessment.  As a community, we need to reflect more on research practices and on how (and when) we reflect on the prominent hypotheses of our field.
\end{abstract}

@@

Accordingly, this paper revisits
the {\em phase delay effect}
i.e.,  the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. Our surveys of industrial practitioners (shown in \S\ref{survey}) reveal the phase-delay effect to be
the single most strongly held belief amongst commercial SE engineers.
Previously~\cite{me08a}, we have constructed return-on-investement models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, who many  unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers,
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier that the developers then, due to the phase delay effect, this saves the overall cost of the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

% Also, as discussed later in this paper, concerns over phase delay were a major driver in the development of agile software processes.

In this paper, we investigate two hypotheses:

\bi
    \item $H_1$: The cost-to-fix of defects monotonically increases across phases.
    \item $H_2$: The cost-to-fix of defects monotonically increases as the phase delay increases.
\ei
The rest of the paper investigates these two hypotheses. First, we conduct a survey of practitioners and researchers to gauge their belief in the phase delay effect. Second, we  analyze 171 software  projects developed in the period \bill{2004 to 2014}.  To the best of our knowledge, this is the largest study on phase delay yet seen in the literature.

% The conclusions of this paper is that it is erroneous to believe in phase delay.
% When we looked for supporting evidence for that effect, we found that:
% \be 
% \item The evidence for that effect in the historical literature is very weak (to say the least);
% \item  In  a large sample of   contemporary software projects, there is no evidence for phase delay.
%  \ee
% The rest of this paper present evidence for the above two points.

% The first point is  documented via  a literature survey and the second point is demonstrated
% via an analysis of

The conclusions of this paper are two fold. First, the phase delay effect does not hold in the large body of software projects analyzed. Second, commonly-held beliefs, such as phase delay, must be periodically revisited empirically in order for researchers to provide accurate and up-to-date guidance for the practitioner community.

% the SE community needs to stop
% justifying changes to SE
% practices via the phase delay effect. Secondly,   myths like phase delay should not have lasted decades
% in the literature without  critical reassessment.  As a community, we need to reflect more on our
% research practices and on how (and when) we reflect and revisit the prominent hypotheses of our field.

@@

Before exploring the commonly-held belief of the {\em phase delay effect}, we digress to discuss the merits of revisiting
old conclusions in software engineering.

Belief in general
principles of software
engineering are common to both research and practice. Processional societies assume such generalities exist when they offer
 lists of supposedly general ``best practices'' such as
the IEEE 1012 standard for software verification~\cite{1012}.  Endres \& Rombach's comment that 
\begin{quote}
Based on repeated and consistent observations, key lessons of these fields can now be formulated into rules or even laws, providing initial building blocks towards a theoretical foundation that is essential for further research, for teaching and for the practice of software development~\cite{endres03,rombach11}
\end{quote}
 Endres \& Rombach offer dozens of lessons of software engineering~\cite{rombach11}.
 Many other 
widely-cited researchers  do the same; e.g.
Glass~\cite{glass02}; Jones~\cite{jones07}; Boehm~\cite{boehm00b}.
Budgen \& Kitchenham seek to reorganize SE research using
general
conclusions drawn from a larger number of studies~\cite{kitch04,budgen09}.

@@

\item
Ray et al.~\cite{ray2014lang} tested if   strongly typed languages
predict for better code quality. In  728 projects,
they found  only a modest benefit in strong typing
(and caution that that effect may actually be due to other  conflating factors).
\item
Fenton \& Neil~\cite{fenton00,fenton00b}   critique the truism that
``pre-release fault rates for software
are a predictor for post-release failures'' (as claimed by~\cite{dunsmore88},
amongst others). For the systems described in~\cite{fenton97}, they
show that software modules that were highly fault-prone
prior to release revealed very few faults after release.
\item
Meyer claims that   object-oriented (OO) encapsulation will
reduce error rates in software~\cite{Meyer1988}.  Yet empirical results suggest
that debugging an OO program is many times harder and
longer than debugging a standard procedural program~\cite{hatton98}.
\item A truism of visual programming is that ``visual
representations are inherently superior to mere textual representations''. A review by Menzies suggests that the available
evidence for this claim is hardly conclusive~\cite{me00v}. 
\item Numerous recent {\em local learning} results compare single models
learned from all available data to multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,betta12,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions).
 \item 
Other work studies  found dozens of single
models that performed worse that {\em committees of models}, all with different opinions,
that contribute to an {\em ensemble} prediction of software development effort~\cite{kocaguneli2012value,azhar13}.
\ei
%To be fair to the software engineering community,
Software engineering is not the only
field where practitioners hang on to persistent beliefs, even if the evidence
for those beliefs is not strong.
The medical profession applies  many practices based on studies
that have been disproved. For example,
a  recent article
in the Mayo Clinic Proceedings~\cite{prasad13} found  
146 medical practices based on studies 
in year $i$, but which were  reversed by subsequent trials within years $i+10$.
Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
Aschwanden warns:
\begin{quote}
Cognitive biases - such as motivated reasoning (all of us want to believe that the things we do make a difference), base rate neglect (failing to pay attention to what happens in the absence of the intervention), and confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) - also influence how we process information~\cite{aschwanden15}.
\end{quote}

The cognitive issues that complicate medicine are also found in software engineering.
According to Passos et al.~\cite{passos11}, commercial developers
are all too willing to believe in general
truisms.
They  warn that developers
usually assume that the truisms they learn from a few past
projects are general to 
all their future projects:
\begin{quote}\label{q:pass}
...the past experiences were taken into account without 
much consideration for their context~\cite{passos11}.  
\end{quote}
The results of J{\o}rgensen \& Gruschke~\cite{jorgensen09} support the findings if Passos et al. They report that 
  supposed software engineering    ``experts'' rarely use lessons
  from past projects to improve their future reasoning~\cite{jorgensen09}. 
 They note that
when the experts
  fail to revise their beliefs, this leads to poor
 conclusions and software projects  (see examples in~\cite{jorgensen09}).
%% This leads to  detrimental
%% situations where groups or 
%% individuals working on the same project have conflicting beliefs, which they never debate
%or refine or unify.

In summary, just like medicine, our field suffers when
 software engineers do  revise old beliefs.  Therefore, it is important
 to regularly  reexamine    old beliefs.

\subsection{Why Study Phase Delay?}
\todo{\begin{itemize}
\item Given the Intro and Motivation, this section should focus on the influence of phase delay in software engineering milieux
\item Need to tone down the reliance on Beck. The important Beck point is that eliminating the cost-to-fix curve (resulting from phase delay) is one of the main goals of agile. His observations on new technologies are good too.
\item Need to cite other popular sources of this information, e.g., McConnell's "Ounce of Prevention" article~\cite{mcconnell01}.
\item Robert Glass's~\cite{glass02} book is very popular and contains an entry on this point
\item Google Scholar Citations for phase delay works Boehm88~\cite{boehm88} (673), Boehm \& Basili~\cite{boehm01} (761), 
\end{itemize}}

Due to its prominence in the field of software engineering,
this paper studies phase delay. As mentioned in the introduction, concerns about phase delay have
been used for decades to motivate major changes to software engineering practices. For example,
in 2000, Kent Beck used phase delay to motivate his call for ``extreme programming'' (a set of techniques
that evolved into contemporary agile and DevOps practices). Beck writes:
\begin{quote}
One of the universal assumptions of software engineering is that the cost of changing a program rises exponentially over time. I can remember sitting in a big linoleum-floored classroom as a college junior and seeing the professor draw on the board the curve found in \fig{curve1}.

@@

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 & **\\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+2$ & 1 & 4& 4 &******** \\
Phase $i$ &Phase $i+3$ & 1 & 8 & 8 & ****************  
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve1}.}\label{fig:curve1a}
\end{figure}

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 & ** \\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+2$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & **** \\
 Phase $i$  &Phase $i+4$ &1 & 2& 2 &**** 
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
\end{figure}

@@

\section{A Survey on Phase Delay and Cost-to-fix Belief}
\label{survey}

@@

%\section{carterCharts}

%\input{carter}

\begin{figure*}[!t]
\begin{center}
\includegraphics[width=6in]{waterfall3.png}
\end{center}
\caption{Different approaches to software development:  waterfall and agile (bottom left).}
\label{fig:waterfall}
\end{figure*}

\section{Data Collection}

\section{Analysis of Phase Delay in 171 TSP Projects}

\todo{
\bi
    \item What is the cost-to-fix curve in TSP data? Check against H1 and compare to past literature.
    \item What is the effect of phase delay? Check against H2.
\ei
}

\subsection{TSP Project Cost-to-fix curve}

The distribution of defects found and fixed per phase is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections. \todo{How does this compare evidence from other sources?}

The median time to fix a defect in each phase is shown in \fig{fix-time-per-phase}. \todo{Overlay median on Boehm's chart \fig{cost-to-fix}. How does it compare?}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{fix-phase-dist.png}
\end{center}
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\begin{tabular}{r|rrr|l}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase removed \\
\hline
120 & 3 & 3 & 18 & ReqtsInspect.dat \\
126 & 3 & 3 & 20 & ReqtsReview.dat \\
70 & 2 & 2 & 10 & HLDInspect.dat \\
 55 &    2 &    1&    7&HLDReview.dat \\

278&    7&    7&   33&DesignInspect.dat \\
233&    6&    6&   25&DesignReview.dat \\

315&    8&    7&   34&CodeInspect.dat \\ 
282&    7&    7&   32&CodeReview.dat\\

425&   11&   10&   54&UnitTest.dat \\ 
226&    7&    6&   46&IntTest.dat \\ 
194&    7&    6&   52&SysTest.dat\\ 
 88&    4&    3&   38&AcceptTest.dat\\
\end{tabular}
\end{center}
\caption{Median fix time per phase}
\label{fig:fix-time-per-phase}
\end{figure}

We reject $H_1$.

\subsection{Phase delay analysis}
\todo{
\bi
    \item What can we say about \fig{raw}
    \item What can we say about \fig{scale}
    \item What is the difference between \fig{raw} and \fig{scale}
\ei
}

@@

We reject $H_2$.

\begin{figure*}[!t] 
 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll|rl}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
\\
  29 &     1 &     3 &    12 & Before & CodeInspect  & 1.00  &  **  \\
  26 &     1 &     4 &     6 & Before & CodeReview  & 1.33  &  ***  \\
  73 &     5 &    13 &    59 & Before & UnitTest  & 4.33  &  *********  \\
  36 &     7 &    24 &    54 & Before & IntTest  & 8.00  &  ****************  \\
  34 &     8 &    28 &    56 & Before & SysTest  & 9.33  &  *******************  \\\hline
\\
128 &     2 &     6 &    15 & Reqts & ReqtsInspect  & 1.00  &  **  \\
132 &     2 &     5 &    12 & Reqts & ReqtsReview  & 0.83  &  **  \\
 28 &     3 &    13 &    29 & Reqts & DesignInspect  & 2.16  &  *****  \\
 22 &     6 &    14 &    31 & Reqts & DesignReview  & 2.33  &  *****  \\\hline
\\
272 &     6 &    14 &    30 & Design & DesignInspect  & 1.00  &  **  \\
 32 &     1 &     4 &    13 & Design & DesignInspection  & 0.28  &  *  \\
240 &     4 &    11 &    22 & Design & DesignReview  & 0.78  &  **  \\
110 &     3 &    11 &    23 & Design & CodeInspect  & 0.78  &  **  \\
 91 &     3 &     7 &    18 & Design & CodeReview  & 0.50  &  *  \\
323 &     8 &    21 &    45 & Design & UnitTest  & 1.50  &  ***  \\
 56 &     7 &    20 &    41 & Design & IntTest  & 1.42  &  ***  \\
 69 &     9 &    27 &    53 & Design & SysTest  & 1.92  &  ****  \\\hline
\\
310 &     6 &    15 &    31 & Code & CodeInspect  & 1.00  &  **  \\
285 &     5 &    14 &    29 & Code & CodeReview  & 0.93  &  **  \\
287 &     7 &    17 &    44 & Code & UnitTest  & 1.13  &  ***  \\
180 &     7 &    20 &    45 & Code & IntTest  & 1.33  &  ***  \\
137 &     6 &    15 &    35 & Code & SysTest  & 1.00  &  **  \\
 37 &     1 &    11 &    27 & Code & AcceptTest  & 0.73  &  **  \\

 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data.}
\label{fig:raw}
\end{figure*}

@@

@@

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{99th }\\
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltasBigBig}
\end{tabular}
\end{center}
\caption{99th percentile scale ups.}
\label{fig:scale99}
\end{figure}
